### 2009-07-14 paper.tex
Fixed tables per plos request.

\begin{table}[!ht]
  \caption{
    \textbf{State types of the singlet transducer (single-sequence
      SCFG) of the TKF Structure Tree model.}}
  \begin{tabular}{|l|rrrl|}
    \hline
    State & $\type$ & $\absorb$ & $\emit$ & description \\ \hline
    $L$ & $\Sstart$ & & & Start of a loop \\ \hline
    $I_L$ & $\Sinsert$ & & $(x,\Tnull)$ & Single-base emission \\ \hline \hline
    $S$ & $\Sstart$ & & & Start of a stem \\ \hline
    $I_S$ & $\Sinsert$ & & $(x,y)$ & Base-pair emission \\ \hline \hline
    $B$ & $\Sinsert$ & & $(L\,S)$ & Bifurcation \\ \hline
  \end{tabular}
  \begin{flushleft}
    Singlet transducers can only have states of type $\Sstart$ or
    $\Sinsert$.
  \end{flushleft}
 \tablabel{tkfstsinglet-types}
\end{table}

% Use table to allow the table to span the width of the entire page.
\begin{table}[!ht]
  \caption{
    \textbf{Singlet transducer (single-sequence SCFG) of the TKF Structure Tree model.}}
  \begin{tabular}{|rcl|r||rcl|r|}
    \hline
    Source & $\rightarrow$ & Destination & probability & Source & $\rightarrow$ & Destination & probability \\ \hline
    $L$ & $\rightarrow$ & $u \,\, I_L$ & $\kappa_l \cdot \pi_l(u)$ & $S$ & $\rightarrow$ & $u \,\, I_S \,\, v$ & $\kappa_s \cdot \pi_s(uv)$ \\ \hline
    & $|$ & $B$ & $\kappa_l \cdot \pi_l(S)$ & & $|$ & $B_e$ & $1-\kappa_s$ \\ \hline
    & $|$ & $\Send$ & $1-\kappa_l$ & & & & \\ \hline
    $I_L$ & $\rightarrow$ & $u \, \, I_L$ & $\kappa_l \cdot \pi_l(u)$ & $I_S$ & $\rightarrow$ & $u \,\, I_S \,\, v$ & $\kappa_s \cdot \pi_s(uv)$\\ \hline
    & $|$ & $B$ & $\kappa_l \cdot \pi_l(S)$ & & $|$ & $B_e$ & $1-\kappa_s$\\ \hline
    & $|$ & $\Send$ & $1-\kappa_l$ & & & & \\ \hline \hline
    $B$ & $\rightarrow$ & $(L \,\, S)$ & 1 & & & & \\ \hline
    $B_e$ & $\rightarrow$ & $(L \,\, \Send)$ & 1 & & & & \\ \hline
  \end{tabular}
  \begin{flushleft}
    The state types for this model are shown in
    \tabref{tkfstsinglet-types}.  The singlet transducer generates
    ancestral RNA sequences and structures.  We use the notation of
    formal grammars to represent state transformation rules; for
    example, the rule $I_L \rightarrow u\,\,I_L$ corresponds to (in a
    Pair HMM) an $\Sinsert$ state $I_L$ emitting a nucleotide $u$ and
    then making a self-transition.  Both loop ($L$ and $I_L$) and stem
    ($S$ and $I_S$) sequence evolve as TKF sequences with length
    parameters $\kappa_l$ and $\kappa_s$ (defined in \secname{A simple model
    of RNA structural evolution}).  $\pi_l(u)$ and $\pi_s(uv)$ are
    the equilibrium distributions of unpaired nucleotides $u$ and
    paired nucleotides $(u,v)$ and are normalized such that
    $\pi_l(S) + \sum_u \pi_l(u) = 1$ and $\sum_{u,v} \pi_s(uv) = 1$.  The bifurcation state $B_e$ is used
    to end stem sequences (only loop sequences are allowed to
    transition to the empty string).
  \end{flushleft}
  \tablabel{tkfstsinglet}
\end{table}

\begin{table}[!ht]
  \caption{
    \textbf{State types of the branch transducer
      (conditionally-normalized Pair SCFG) of the TKF Structure Tree
      model.}}
  \begin{tabular}{|l|rrrl|} \hline
    State & $\type$ & $\absorb$ & $\emit$ & description \\ \hline
    $L$ & $\Sstart$ & & & Start of a loop \\ \hline
    $I_L$ & $\Sinsert$ & & $(u,\Tnull)$ & Single-base insertion \\ \hline
    $M_L$ & $\Smatch$ & $(x,\Tnull)$ & $(u,\Tnull)$ & Single-base substitution \\ \hline
    $D_L$ & $\Smatch$ & $(x,\Tnull)$ & & Single-base deletion \\ \hline
    $W_L$ & $\Swait$ & & & Wait for next base \\ \hline \hline
    $S$ & $\Sstart$ & & & Start of a stem \\ \hline
    $I_S$ & $\Sinsert$ & & $(u,v)$ & Base-pair insertion \\ \hline
    $M_S$ & $\Smatch$ & $(x,y)$ & $(u,v)$ & Base-pair substitution \\ \hline
    $D_S$ & $\Smatch$ & $(x,y)$ & & Base-pair deletion \\ \hline
    $W_S$ & $\Swait$ & & & Wait for next base-pair \\ \hline \hline
    $B_i$ & $\Sinsert$ & & $(L_i\,S_i)$ & Stem insertion \\ \hline
    $B$ & $\Smatch$ & $(L\,S)$ & $(L\,S)$ & Stem conservation \\ \hline
    $B_p$ & $\Smatch$ & $(L\,S)$ & $(L\,\Send)$ & Stem deletion \\ \hline
    $B_e$ & $\Smatch$ & $(L\,\Send)$ & $(L\,\Send)$ & Stem extinction \\ \hline
  \end{tabular}
  \begin{flushleft}
    States which have the same names as states of the singlet transducer in \tabref{tkfstsinglet-types}
    are the branch-transducer equivalents of the corresponding singlet-transducer states
    (e.g., a $\Smatch$ state might be the branch equivalent of an $\Sinsert$ state).
    States $L_i$ and $S_i$ are the $\Sstart$ states of a sub-model (not shown) identical
    in structure to the singlet transducer.  They are used to insert a
    new stem-loop structure.
  \end{flushleft}
  \tablabel{tkfstbranch-types}
\end{table}

% Use table to allow the table to span the width of the entire page.
\begin{table}[!ht]
  \caption{
    \textbf{Branch transducer (conditionally-normalized Pair SCFG) of
      the TKF Structure Tree model.}}
  \begin{tabular}{|rcl|r||rcl|r|} \hline
    Source & $\rightarrow$ & Destination & probability & Source & $\rightarrow$ & Destination & probability \\ \hline
    $L$ & $\rightarrow$ & $w \,\, I_L$ & $\beta_l(t) \cdot \pi_l(w)$ & $S$ & $\rightarrow$ & $w \,\, I_S \,\, x$ & $\beta_s(t) \cdot \pi_s(wx)$ \\ \hline
    & $|$ & $B_i$ & $\beta_l(t) \cdot \pi_l(S)$ & & $|$ & $W_S$ & $1-\beta_s(t)$ \\ \hline
    & $|$ & $W_L$ & $1-\beta_l(t)$ & & & &  \\ \hline
    $I_L$ & $\rightarrow$ & $w \,\, I_L$ & $\beta_l(t) \cdot \pi_l(w)$ & $I_S$ & $\rightarrow$ & $w \,\, I_S \,\, x$ & $\beta_s(t) \cdot \pi_s(wx)$ \\ \hline
    & $|$ & $B_i$ & $\beta_l(t) \cdot \pi_l(S)$ & & $|$ & $W_S$ & $1-\beta_s(t)$ \\ \hline
    & $|$ & $W_L$ & $1-\beta_l(t)$ & & & & \\ \hline
    $M_L$ & $\rightarrow$ & $w \,\, I_L$ & $\beta_l(t) \cdot \pi_l(w)$ & $M_S$ & $\rightarrow$ & $w \,\, I_S \,\, x$ & $\beta_s(t) \cdot \pi_s(wx)$ \\ \hline
    & $|$ & $B_i$ & $\beta_l(t) \cdot \pi_l(S)$ & & $|$ & $W_S$ & $1-\beta_s(t)$ \\ \hline
    & $|$ & $W_L$ & $1-\beta_l(t)$ & & & &  \\ \hline
    $D_L$ & $\rightarrow$ & $w \,\, I_L$ & $\gamma_l(t) \cdot \pi_l(w)$ & $D_S$ & $\rightarrow$ & $w \,\, I_S \,\, x$ & $\gamma_s(t) \cdot \pi_s(wx)$ \\ \hline
    & $|$ & $B_i$ & $\gamma_l(t) \cdot \pi_l(S)$ & & $|$ & $W_S$ & $1-\gamma_s(t)$ \\ \hline
    & $|$ & $W_L$ & $1-\gamma_l(t)$ & & & &  \\ \hline
    $W_L$ & $\rightarrow$ & $w \,\, M_L$ & $\alpha_l(t) \cdot M_l(u \rightarrow w)$ & $W_S$ & $\rightarrow$ & $w \,\, M_S \,\, x$ & $\alpha_s(t) \cdot M_s(uv \rightarrow wx)$ \\ \hline
    & $|$ & $D_L$ & $1-\alpha_l(t)$ & & $|$ & $D_S$ & $1-\alpha_s(t)$ \\ \hline
    & $|$ & $B$ & $\alpha_l(t)$ & & $|$ & $B_e$ & $1$ \\ \hline
    & $|$ & $B_p$ & $1-\alpha_l(t)$ & & & &  \\ \hline
    & $|$ & $\Send$ & $1$ & & & &  \\ \hline \hline
    $B$ & $\rightarrow$ & $L \,\, S$ & 1 & $B_p$ & $\rightarrow$ & $L \, \, \Send$ & 1 \\ \hline
    $B_i$ & $\rightarrow$ & $L_i \,\, S_i$ & 1 & $B_e$ & $\rightarrow$ & $L \, \, \Send$ & 1 \\ \hline \hline
  \end{tabular}
  \begin{flushleft}
    The state types for this model are shown in
    \tabref{tkfstbranch-types}. The branch transducer evolves a
    sequence and structure along a branch of the phylogenetic
    tree. States $L_i$ and $S_i$ are the $\Sstart$ states for a
    sub-model corresponding to an insertion of a new stem in the
    descendant sequences; the sub-model (not shown) is identical in
    structure to the singlet transducer shown in
    \tabref{tkfstsinglet}. $\pi_l(w)$ and $\pi_s(wx)$ are the
    equilibrium distributions of, respectively, descendant unpaired
    nucleotide $w$ and descendant paired nucleotides $(w,x)$;
    $M_l(u \rightarrow w)$ and $M_l(uv \rightarrow wx)$ are the conditional distributions
    (i.e., match probabilities) of a descendant unpaired
    nucleotide $w$ given an ancestral unpaired nucleotide $u$ and
    descendant paired nucleotides $(w,x)$ given ancestral nucleotides
    $(u,v)$. The functions $\alpha_{l,s}(t)$, $\beta_{l,s}(t)$ and
    $\gamma_{l,s}(t)$ are parametrized by the insertion and deletion rates
    of the TKFST model and are defined in \secname{A simple model of RNA
    structural evolution}.
  \end{flushleft}
  \tablabel{tkfstbranch}
\end{table}

@@

\begin{table}[ht]
  \caption{
    \textbf{Percentage sensitivity and positive predictive value (Sensitivity / PPV) for pairwise nucleotide-level alignments in the \bralibaseII\ benchmark.}}
  \begin{tabular}{|r|rrrr|}
    \hline
    & U5                      & g2intron                & rRNA                    & tRNA \\
    \hline
    TKFST grammar     &      81.6  /      81.7  & {\bf 75.4} / {\bf
      75.0} &      91.4  /      92.6  & {\bf 94.6} / {\bf 94.4} \\ \hline
    \stemloc\ grammar & {\bf 82.6} / {\bf 83.7} &      74.2  /
    74.8  & {\bf 92.6} / {\bf 92.8} &      93.2  /      93.9 \\ \hline
    \end{tabular}
\begin{flushleft}
  We compared the performance of the TKFST model for progressive multiple alignment
  of RNAs against the performance of a grammar with a richer model of RNA structure (\stemloc\ \cite{Holmes2005}).
  Sensitivity is defined as $\mbox{TP}/(\mbox{TP} + \mbox{FN})$ and PPV is defined as $\mbox{TP}/(\mbox{TP} + \mbox{FP})$,
  where TP is the number of true positives (correctly aligned residue pairs),
  FN is the number of false negatives (residue pairs that should have been aligned but were not) and
  FP is the number of false positives (residue pairs that were incorrectly aligned).
  These statistics are summed over all pairs of sequences in the multiple alignment;
  therefore, ``Sensitivity'' for pairwise residue alignments is equivalent to the Sum of Pairs Score or SPS \cite{ThompsonEtal99}.
  ``g2intron'' is the RFAM entry {\tt Intron\_gpII}, containing domains V and VI of the Group II intron.
  \end{flushleft}
  \tablabel{bralibaseII-align}
\end{table}

\begin{table}[ht]
  \caption{
    \textbf{Percentage sensitivity and positive predictive value (Sensitivity / PPV) for predicted base-pairs in the \bralibaseII\ benchmark.}}
  \begin{tabular}{|r|rrrr|}
    \hline
    & U5                      & g2intron                & rRNA                    & tRNA \\
    \hline
    TKFST grammar      &      37.9  /      68.0   &      42.1  /
    63.8  &      37.4  /      66.5  &      70.9  / {\bf 88.3} \\ \hline
    \stemloc\ grammar  &      74.9  / {\bf 73.9}  &      64.3  /
    56.7  &      51.0  /      59.0  &      74.0  /      76.4  \\ \hline
  \end{tabular}
  \begin{flushleft}
    We compared the performance of the TKFST model for structure-prediction accuracy during
    progressive multiple alignment of RNAs against the performance of a grammar with 
    a richer model of RNA structure (\stemloc\ \cite{Holmes2005}).
    ``g2intron'' is the RFAM entry {\tt Intron\_gpII}, containing domains V and VI of the Group II intron.
  \end{flushleft}
  \tablabel{bralibaseII-struct}
\end{table}

% nanos: Counted 3672701 subseqs and 48323860 bifurcations.
% tRNA: Counted 12573792 subseqs and 404730480 bifurcations.
% Y: Counted 34474759 subseqs and 1429713088 bifurcations.
% g2intron: Counted 139037184 subseqs and 1895528744 bifurcations.
% Scaling constant is approximately
% A * (404730480 / 48323860) = (19 / 3) => A = 0.76
% time = A * (# bifs / # bifs for nanos) * time for nanos
% Y time = 0.76 * (1429713088 / 48323860) * 3 min = 67 min
% g2intron time = 0.76 * (1895528744 / 48323860) * 3 min = 89 min
\begin{table}[!ht]
  \caption{
   \textbf{Estimates of the memory and time required to reconstruct ancestral
     structures of three RNAs from several families of biological interest
     (as reported by \indiegram).}}
 \begin{tabular}{|l|rrr|} \hline
    Family & Sequence lengths & Memory & Time \\ \hline
    \emph{nanos} 3' TCE & 61-64 nt & 3 Gb & 3 min \\ \hline
    tRNA & 69-73 nt & 11 Gb & 19 min \\ \hline
    Y RNA & 47-81 nt & 33 Gb & 70 min \\ \hline
    Group II intron (domains V and VI) & 76-91 nt & 122 Gb & 90 min \\ \hline
  \end{tabular}
  \begin{flushleft}
    The \emph{nanos} 3' translational control element (TCE) sequences
    are the seed sequences of the
    corresponding RFAM family \cite{GriffithsJonesEtAl2003} and the
    three tRNA sequences are from the \bralibaseII\ database
    \cite{GardnerEtAl2005} (identifiers AB042432.1-14140\_14072,
    Z82044.1-16031\_16103 and AC008670.6-83725\_83795). The group II
    intron sequences (identifiers Z00044.1-87253\_87177,
    X57546.1-2817\_2907 and X04465.1-2700\_2775) are from BralibaseII
    \cite{GardnerEtAl2005}.  The Y RNAs are hY1, hY4, and hY5 from
    \cite{TeunissenEtAl2000}; sequence lengths exclude the conserved
    stem S1.  The time estimates are for a 2.2GHz AMD Opteron 848 CPU.
  \end{flushleft}
  \tablabel{complexities}
\end{table}

>>>>

\begin{table}[!ht]
  \caption{
    \textbf{State types of the singlet transducer (single-sequence
      SCFG) of the TKF Structure Tree model.}}
  \begin{tabular}{|l|r|r|r|l|}
    \hline
    State & $\type$ & $\absorb$ & $\emit$ & description \\ \hline
    $L$ & $\Sstart$ & & & Start of a loop \\ \hline
    $I_L$ & $\Sinsert$ & & $(x,\Tnull)$ & Single-base emission \\ \hline \hline
    $S$ & $\Sstart$ & & & Start of a stem \\ \hline
    $I_S$ & $\Sinsert$ & & $(x,y)$ & Base-pair emission \\ \hline \hline
    $B$ & $\Sinsert$ & & $(L\,S)$ & Bifurcation \\ \hline
  \end{tabular}
  \begin{flushleft}
    Singlet transducers can only have states of type $\Sstart$ or
    $\Sinsert$.
  \end{flushleft}
 \tablabel{tkfstsinglet-types}
\end{table}

% Use table to allow the table to span the width of the entire page.
\begin{table}[!ht]
  \caption{
    \textbf{Singlet transducer (single-sequence SCFG) of the TKF Structure Tree model.}}
  \begin{tabular}{|l|r||l|r|}
    \hline
    Source $\rightarrow$ Destination & probability & Source $\rightarrow$ Destination & probability \\ \hline
    $L \rightarrow u \,\, I_L$ & $\kappa_l \cdot \pi_l(u)$ & $S \rightarrow u \,\, I_S \,\, v$ & $\kappa_s \cdot \pi_s(uv)$ \\ \hline
    $\phantom{L} \rightarrow B$ & $\kappa_l \cdot \pi_l(S)$ & $\phantom{S} \rightarrow B_e$ & $1-\kappa_s$ \\ \hline
    $\phantom{L} \rightarrow \Send$ & $1-\kappa_l$ & & \\ \hline
    $I_L \rightarrow u \, \, I_L$ & $\kappa_l \cdot \pi_l(u)$ & $I_S \rightarrow u \,\, I_S \,\, v$ & $\kappa_s \cdot \pi_s(uv)$\\ \hline
    $\phantom{I_L} \rightarrow B$ & $\kappa_l \cdot \pi_l(S)$ & $\phantom{I_S} \rightarrow B_e$ & $1-\kappa_s$\\ \hline
    $\phantom{I_L} \rightarrow \Send$ & $1-\kappa_l$ & & \\ \hline \hline
    $B \rightarrow (L \,\, S)$ & 1 & & \\ \hline
    $B_e \rightarrow (L \,\, \Send)$ & 1 & & \\ \hline
  \end{tabular}
  \begin{flushleft}
    The state types for this model are shown in
    \tabref{tkfstsinglet-types}.  The singlet transducer generates
    ancestral RNA sequences and structures.  We use the notation of
    formal grammars to represent state transformation rules; for
    example, the rule $I_L \rightarrow u\,\,I_L$ corresponds to (in a
    Pair HMM) an $\Sinsert$ state $I_L$ emitting a nucleotide $u$ and
    then making a self-transition.  Both loop ($L$ and $I_L$) and stem
    ($S$ and $I_S$) sequence evolve as TKF sequences with length
    parameters $\kappa_l$ and $\kappa_s$ (defined in \secname{A simple model
    of RNA structural evolution}).  $\pi_l(u)$ and $\pi_s(uv)$ are
    the equilibrium distributions of unpaired nucleotides $u$ and
    paired nucleotides $(u,v)$ and are normalized such that
    $\pi_l(S) + \sum_u \pi_l(u) = 1$ and $\sum_{u,v} \pi_s(uv) = 1$.  The bifurcation state $B_e$ is used
    to end stem sequences (only loop sequences are allowed to
    transition to the empty string).
  \end{flushleft}
  \tablabel{tkfstsinglet}
\end{table}

\begin{table}[!ht]
  \caption{
    \textbf{State types of the branch transducer
      (conditionally-normalized Pair SCFG) of the TKF Structure Tree
      model.}}
  \begin{tabular}{|l|r|r|r|l|} \hline
    State & $\type$ & $\absorb$ & $\emit$ & description \\ \hline
    $L$ & $\Sstart$ & & & Start of a loop \\ \hline
    $I_L$ & $\Sinsert$ & & $(u,\Tnull)$ & Single-base insertion \\ \hline
    $M_L$ & $\Smatch$ & $(x,\Tnull)$ & $(u,\Tnull)$ & Single-base substitution \\ \hline
    $D_L$ & $\Smatch$ & $(x,\Tnull)$ & & Single-base deletion \\ \hline
    $W_L$ & $\Swait$ & & & Wait for next base \\ \hline \hline
    $S$ & $\Sstart$ & & & Start of a stem \\ \hline
    $I_S$ & $\Sinsert$ & & $(u,v)$ & Base-pair insertion \\ \hline
    $M_S$ & $\Smatch$ & $(x,y)$ & $(u,v)$ & Base-pair substitution \\ \hline
    $D_S$ & $\Smatch$ & $(x,y)$ & & Base-pair deletion \\ \hline
    $W_S$ & $\Swait$ & & & Wait for next base-pair \\ \hline \hline
    $B_i$ & $\Sinsert$ & & $(L_i\,S_i)$ & Stem insertion \\ \hline
    $B$ & $\Smatch$ & $(L\,S)$ & $(L\,S)$ & Stem conservation \\ \hline
    $B_p$ & $\Smatch$ & $(L\,S)$ & $(L\,\Send)$ & Stem deletion \\ \hline
    $B_e$ & $\Smatch$ & $(L\,\Send)$ & $(L\,\Send)$ & Stem extinction \\ \hline
  \end{tabular}
  \begin{flushleft}
    States which have the same names as states of the singlet transducer in \tabref{tkfstsinglet-types}
    are the branch-transducer equivalents of the corresponding singlet-transducer states
    (e.g., a $\Smatch$ state might be the branch equivalent of an $\Sinsert$ state).
    States $L_i$ and $S_i$ are the $\Sstart$ states of a sub-model (not shown) identical
    in structure to the singlet transducer.  They are used to insert a
    new stem-loop structure.
  \end{flushleft}
  \tablabel{tkfstbranch-types}
\end{table}

% Use table to allow the table to span the width of the entire page.
\begin{table}[!ht]
  \caption{
    \textbf{Branch transducer (conditionally-normalized Pair SCFG) of
      the TKF Structure Tree model.}}
  \begin{tabular}{|l|r||l|r|} \hline
    Source $\rightarrow$ Destination & probability & Source $\rightarrow$ Destination & probability \\ \hline
    $L \rightarrow w \,\, I_L$ & $\beta_l(t) \cdot \pi_l(w)$ & $S \rightarrow w \,\, I_S \,\, x$ & $\beta_s(t) \cdot \pi_s(wx)$ \\ \hline
    $\phantom{L} \rightarrow B_i$ & $\beta_l(t) \cdot \pi_l(S)$ & $\phantom{S} \rightarrow W_S$ & $1-\beta_s(t)$ \\ \hline
    $\phantom{L} \rightarrow W_L$ & $1-\beta_l(t)$ & & \\ \hline
    $I_L \rightarrow w \,\, I_L$ & $\beta_l(t) \cdot \pi_l(w)$ & $I_S \rightarrow w \,\, I_S \,\, x$ & $\beta_s(t) \cdot \pi_s(wx)$ \\ \hline
    $\phantom{I_L} \rightarrow B_i$ & $\beta_l(t) \cdot \pi_l(S)$ & $\phantom{I_S} \rightarrow W_S$ & $1-\beta_s(t)$ \\ \hline
    $\phantom{I_L} \rightarrow W_L$ & $1-\beta_l(t)$ & & \\ \hline
    $M_L \rightarrow w \,\, I_L$ & $\beta_l(t) \cdot \pi_l(w)$ & $M_S \rightarrow w \,\, I_S \,\, x$ & $\beta_s(t) \cdot \pi_s(wx)$ \\ \hline
    $\phantom{M_L} \rightarrow B_i$ & $\beta_l(t) \cdot \pi_l(S)$ & $\phantom{M_S} \rightarrow W_S$ & $1-\beta_s(t)$ \\ \hline
    $\phantom{M_L} \rightarrow W_L$ & $1-\beta_l(t)$ & & \\ \hline
    $D_L \rightarrow w \,\, I_L$ & $\gamma_l(t) \cdot \pi_l(w)$ & $D_S \rightarrow w \,\, I_S \,\, x$ & $\gamma_s(t) \cdot \pi_s(wx)$ \\ \hline
    $\phantom{D_L} \rightarrow B_i$ & $\gamma_l(t) \cdot \pi_l(S)$ & $\phantom{D_S} \rightarrow W_S$ & $1-\gamma_s(t)$ \\ \hline
    $\phantom{D_L} \rightarrow W_L$ & $1-\gamma_l(t)$ & & \\ \hline
    $W_L \rightarrow w \,\, M_L$ & $\alpha_l(t) \cdot M_l(u \rightarrow w)$ & $W_S \rightarrow w \,\, M_S \,\, x$ & $\alpha_s(t) \cdot M_s(uv \rightarrow wx)$ \\ \hline
    $\phantom{W_L} \rightarrow D_L$ & $1-\alpha_l(t)$ & $\phantom{W_S} \rightarrow D_S$ & $1-\alpha_s(t)$ \\ \hline
    $\phantom{W_L} \rightarrow B$ & $\alpha_l(t)$ & $\phantom{W_S} \rightarrow B_e$ & $1$ \\ \hline
    $\phantom{W_L} \rightarrow B_p$ & $1-\alpha_l(t)$ & & \\ \hline
    $\phantom{W_L} \rightarrow \Send$ & $1$ & & \\ \hline \hline
    $B \rightarrow (L \,\, S)$ & 1 & $B_p \rightarrow (L \, \, \Send)$ & 1 \\ \hline
    $B_i \rightarrow (L_i \,\, S_i)$ & 1 & $B_e \rightarrow (L \, \, \Send)$ & 1 \\ \hline \hline
  \end{tabular}
  \begin{flushleft}
    The state types for this model are shown in
    \tabref{tkfstbranch-types}. The branch transducer evolves a
    sequence and structure along a branch of the phylogenetic
    tree. States $L_i$ and $S_i$ are the $\Sstart$ states for a
    sub-model corresponding to an insertion of a new stem in the
    descendant sequences; the sub-model (not shown) is identical in
    structure to the singlet transducer shown in
    \tabref{tkfstsinglet}. $\pi_l(w)$ and $\pi_s(wx)$ are the
    equilibrium distributions of, respectively, descendant unpaired
    nucleotide $w$ and descendant paired nucleotides $(w,x)$;
    $M_l(u \rightarrow w)$ and $M_l(uv \rightarrow wx)$ are the conditional distributions
    (i.e., match probabilities) of a descendant unpaired
    nucleotide $w$ given an ancestral unpaired nucleotide $u$ and
    descendant paired nucleotides $(w,x)$ given ancestral nucleotides
    $(u,v)$. The functions $\alpha_{l,s}(t)$, $\beta_{l,s}(t)$ and
    $\gamma_{l,s}(t)$ are parametrized by the insertion and deletion rates
    of the TKFST model and are defined in \secname{A simple model of RNA
    structural evolution}.
  \end{flushleft}
  \tablabel{tkfstbranch}
\end{table}

@@

\begin{table}[ht]
  \caption{
    \textbf{Percentage sensitivity and positive predictive value (Sensitivity / PPV) for pairwise nucleotide-level alignments in the \bralibaseII\ benchmark.}}
  \begin{tabular}{|r|r|r|r|r|}
    \hline
    & U5                      & g2intron                & rRNA                    & tRNA \\
    \hline
    TKFST grammar     &      81.6  /      81.7  & {\bf 75.4} / {\bf
      75.0} &      91.4  /      92.6  & {\bf 94.6} / {\bf 94.4} \\ \hline
    \stemloc\ grammar & {\bf 82.6} / {\bf 83.7} &      74.2  /
    74.8  & {\bf 92.6} / {\bf 92.8} &      93.2  /      93.9 \\ \hline
    \end{tabular}
\begin{flushleft}
  We compared the performance of the TKFST model for progressive multiple alignment
  of RNAs against the performance of a grammar with a richer model of RNA structure (\stemloc\ \cite{Holmes2005}).
  Sensitivity is defined as $\mbox{TP}/(\mbox{TP} + \mbox{FN})$ and PPV is defined as $\mbox{TP}/(\mbox{TP} + \mbox{FP})$,
  where TP is the number of true positives (correctly aligned residue pairs),
  FN is the number of false negatives (residue pairs that should have been aligned but were not) and
  FP is the number of false positives (residue pairs that were incorrectly aligned).
  These statistics are summed over all pairs of sequences in the multiple alignment;
  therefore, ``Sensitivity'' for pairwise residue alignments is equivalent to the Sum of Pairs Score or SPS \cite{ThompsonEtal99}.
  ``g2intron'' is the RFAM entry {\tt Intron\_gpII}, containing domains V and VI of the Group II intron.
  \end{flushleft}
  \tablabel{bralibaseII-align}
\end{table}

\begin{table}[ht]
  \caption{
    \textbf{Percentage sensitivity and positive predictive value (Sensitivity / PPV) for predicted base-pairs in the \bralibaseII\ benchmark.}}
  \begin{tabular}{|r|r|r|r|r|}
    \hline
    & U5                      & g2intron                & rRNA                    & tRNA \\
    \hline
    TKFST grammar      &      37.9  /      68.0   &      42.1  /
    63.8  &      37.4  /      66.5  &      70.9  / {\bf 88.3} \\ \hline
    \stemloc\ grammar  &      74.9  / {\bf 73.9}  &      64.3  /
    56.7  &      51.0  /      59.0  &      74.0  /      76.4  \\ \hline
  \end{tabular}
  \begin{flushleft}
    We compared the performance of the TKFST model for structure-prediction accuracy during
    progressive multiple alignment of RNAs against the performance of a grammar with 
    a richer model of RNA structure (\stemloc\ \cite{Holmes2005}).
    ``g2intron'' is the RFAM entry {\tt Intron\_gpII}, containing domains V and VI of the Group II intron.
  \end{flushleft}
  \tablabel{bralibaseII-struct}
\end{table}

% nanos: Counted 3672701 subseqs and 48323860 bifurcations.
% tRNA: Counted 12573792 subseqs and 404730480 bifurcations.
% Y: Counted 34474759 subseqs and 1429713088 bifurcations.
% g2intron: Counted 139037184 subseqs and 1895528744 bifurcations.
% Scaling constant is approximately
% A * (404730480 / 48323860) = (19 / 3) => A = 0.76
% time = A * (# bifs / # bifs for nanos) * time for nanos
% Y time = 0.76 * (1429713088 / 48323860) * 3 min = 67 min
% g2intron time = 0.76 * (1895528744 / 48323860) * 3 min = 89 min
\begin{table}[!ht]
  \caption{
   \textbf{Estimates of the memory and time required to reconstruct ancestral
     structures of three RNAs from several families of biological interest
     (as reported by \indiegram).}}
 \begin{tabular}{|l|r|r|r|} \hline
    Family & Sequence lengths & Memory & Time \\ \hline
    \emph{nanos} 3' TCE & 61-64 nt & 3 Gb & 3 min \\ \hline
    tRNA & 69-73 nt & 11 Gb & 19 min \\ \hline
    Y RNA & 47-81 nt & 33 Gb & 70 min \\ \hline
    Group II intron (domains V and VI) & 76-91 nt & 122 Gb & 90 min \\ \hline
  \end{tabular}
  \begin{flushleft}
    The \emph{nanos} 3' translational control element (TCE) sequences
    are the seed sequences of the
    corresponding RFAM family \cite{GriffithsJonesEtAl2003} and the
    three tRNA sequences are from the \bralibaseII\ database
    \cite{GardnerEtAl2005} (identifiers AB042432.1-14140\_14072,
    Z82044.1-16031\_16103 and AC008670.6-83725\_83795). The group II
    intron sequences (identifiers Z00044.1-87253\_87177,
    X57546.1-2817\_2907 and X04465.1-2700\_2775) are from BralibaseII
    \cite{GardnerEtAl2005}.  The Y RNAs are hY1, hY4, and hY5 from
    \cite{TeunissenEtAl2000}; sequence lengths exclude the conserved
    stem S1.  The time estimates are for a 2.2GHz AMD Opteron 848 CPU.
  \end{flushleft}
  \tablabel{complexities}
\end{table}
